The goal of {apyramid} is to provide a quick method for visualizing census data
stratified by age and one or two categorical variables (e.g. gender and health
status). This is a product of the R4EPIs project; learn more at <https://r4epis.netlify.app/>.

## Installation

You can install {apyramid} from CRAN:

``` r
install.packages("apyramid")
```

<details>
<!--
NOTE: everything inside the details tag will be collapsed and effectively
hidden from the user
-->
<summary style='text-decoration: underline'>Click here for alternative installation options</summary>

If there is a bugfix or feature that is not yet on CRAN, you can install it via
the {drat} package:

```{r eval = FALSE}
# install.packages("drat")
drat::addRepo("R4EPI")
install.packages("apyramid")
```

You can also install the in-development version from GitHub using the {remotes}
package (but there's no guarantee that it will be stable):

```{r eval = FALSE}
# install.packages("remotes")
remotes::install_github("R4EPI/apyramid") 
```

</details>

-------

The {apyramid} package was primarily designed for quick visualisation of
un-aggregated linelist data in field epidemiological situations. It has one
available function:

 - `age_pyramid()` returns age pyramid visualizations of linelist, survey, or
   pre-aggregated census data as a `ggplot` object.


```{r load_packages}
library("apyramid")
library("ggplot2")   # load ggplot2 to control plot aesthetics
library("outbreaks") # load the outbreaks package for linelist data
old_theme <- theme_set(theme_classic(base_size = 18))
```


